---
title: "p8105_hw5_kdt2119"
author: "Kelvin Delali Tamakloe"
date: 11/19/21
output: github_document
---

#Loading all libraries and setting themes, color schemes, and graph output sizes 

```{r, echo = FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(haven)
library(ggplot2)
library(patchwork)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Question 1:


```{r}

#Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”

homicide_df = 
    read_csv("homicide-data.csv") %>% 
    mutate(
      city_state = str_c(city, state, sep = "_"),
      resolved = case_when(
        disposition == "Closed without arrest" ~ "unsolved",
        disposition == "Open/No arrest" ~ "unsolved",
        disposition == "Closed by arrest" ~ "solved"
        )
    ) %>% select(city_state,resolved) %>% 
  filter(city_state != "Tulsa_AL")

#To find the total number of homicides:

total_hom = 
  homicide_df %>% 
  group_by(city_state) %>% 
  summarize(
    homicide_total = n(),
    homicide_unsolved = sum(resolved == "unsolved")
  )
total_hom

#For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

baltimore_df = total_hom %>% 
  filter(city_state == "Baltimore_MD")  
proportion = prop.test(
    x = baltimore_df %>% pull(homicide_unsolved),
    n = baltimore_df %>% pull(homicide_total))

#The sample estimates: 0.6455607; The 95% CI: (0.6276,0.6632)

proportion %>%
  broom::tidy() 

#Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

allcities_df = 
  total_hom %>% 
  mutate(
    prop_tests = map2(.x = homicide_unsolved, .y = homicide_total, ~prop.test(x = .x, n = .y)),
    tidy_tests = map(.x = prop_tests, ~broom::tidy(.x))
  ) %>% 
  select(-prop_tests) %>% 
  unnest(tidy_tests) %>% 
  select(city_state, estimate, conf.low, conf.high)
allcities_df

#Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

allcities_df %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

The raw data from the Washington Post homicide dataset includes `r nrow(homicide_df)` rows and `r ncol(homicide_df)` columns and is focused on unsolved homicides throughout the United States. The dataset includes the victims' first and last names, their race, age, sex, the reported date of the homicide, city and state in which the homicide occurred, as well as a unique ID. The data is already relatively clean.

The dataset has records for 52,129 homicides from 50 cities in the United States. The dataset has 12 columns and includes variables such as victim_race, victim_last, victim_first, victim_age, victim_sex, reported_date and uid. 

Post data cleaning and creation of the city_state variable, there are only 2 variables and a total of 52,128 observations.The total number of homicides and the total number of uncilved homicides in each city_state is contained in the tital_hom data frame.

Baltimore has a sample estimate of 0.6455607. The 95% CI has an upper bound of 0.6276 and a lower bound of 0.6632.


## Question 2


Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time. 1) Start with a dataframe containing all file names; 2) Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe, 3) Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary

```{r}

#To load the data from the relevant file in the repository, and merge spreadsheet to create list.

hw5data = 
  tibble(
    path = list.files("hw5_data")) %>% 
  mutate(
    path = str_c("hw5_data/", path), 
    data = map(path, read_csv) 
     ) %>% 
  unnest(data) %>%
  mutate(
    path = str_replace(path, "hw5_data/", " "), #
    path = str_replace(path, ".csv", " ")
    ) %>%
  separate(path, into = c("group","Subject_ID"),sep = "_")

hw5data_ =
  hw5data %>% 
  pivot_longer(
    week_1:week_8,
    names_to = "Week", 
    names_prefix = "week_", 
    values_to = "Observations"
  ) %>%
  mutate(
    group = str_replace(group, "exp", "Experimental Arm"),
    group = str_replace(group, "con", "Control Arm"),
  ) %>%
  relocate(Subject_ID)

#To make a spaghetti plot showing observations for each subject over time

hw5data_ %>% 
  mutate(Week = as.numeric(Week)) %>%
  ggplot(aes(x = Week, y = Observations, color = group)) +
  geom_path(aes(group = Subject_ID)) +
  geom_point() +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(
      title = "Observations for each Subject Over Time",
      x = "Week",
      y = "Observations"
      )
```

The charts above show observations over time as delineated by study arm. It can be noted that subjects in the control arm appear to report largely the same observation over time while subjects in the experimental arm appear to report a higher number of observations as time goes on. 


## Question 3


The code chunk below loads the iris dataset from the tidyverse package and introduces some missing values in each column. The purpose of this problem is to fill in those missing values. For numeric variables, you should fill in missing values with the mean of non-missing values. For character variables, you should fill in missing values with "virginica". Write a function that takes a vector as an argument; replaces missing values using the rules defined above; and returns the resulting vector. Apply this function to the columns of iris_with_missing using a map statement.

```{r}

#To load the iris dataset for purposes of exploration:

set.seed(10)
iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))

#The function below replaces the NA's in the dataset

jf_iris = function(x) {
   if (is.numeric(x)) { 
   replace_na(x, mean(x, na.rm = TRUE))
   }
   else if (is.character(x)) 
     {replace_na(x, "virginica")}
}
iris_completed = map(iris_with_missing, jf_iris)

#To have a final look to rule out any remaining NA's in the dataset

is.na(iris_completed)
```