---
title: "p8105_hw5_kdt2119"
author: "Kelvin Delali Tamakloe"
date: 11/19/21
output: github_document
---

#Loading all libraries and setting themes, color schemes, and graph output sizes 

```{r, echo = FALSE}
library(tidyverse)
library(dplyr)
library(readxl)
library(haven)
library(ggplot2)
library(patchwork)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Question 1:


```{r}

#Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”

homicide_df = 
    read_csv("homicide-data.csv") %>% 
    mutate(
      city_state = str_c(city, state, sep = "_"),
      resolved = case_when(
        disposition == "Closed without arrest" ~ "unsolved",
        disposition == "Open/No arrest" ~ "unsolved",
        disposition == "Closed by arrest" ~ "solved"
        )
    ) %>% select(city_state,resolved) %>% 
  filter(city_state != "Tulsa_AL")

#To find the total number of homicides:

total_hom = 
  homicide_df %>% 
  group_by(city_state) %>% 
  summarize(
    homicide_total = n(),
    homicide_unsolved = sum(resolved == "unsolved")
  )
total_hom

#For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

baltimore_df = total_hom %>% 
  filter(city_state == "Baltimore_MD")  
proportion = prop.test(
    x = baltimore_df %>% pull(homicide_unsolved),
    n = baltimore_df %>% pull(homicide_total))

#The sample estimates: 0.6455607; The 95% CI: (0.6276,0.6632)

proportion %>%
  broom::tidy() 

#Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

allcities_df = 
  total_hom %>% 
  mutate(
    prop_tests = map2(.x = homicide_unsolved, .y = homicide_total, ~prop.test(x = .x, n = .y)),
    tidy_tests = map(.x = prop_tests, ~broom::tidy(.x))
  ) %>% 
  select(-prop_tests) %>% 
  unnest(tidy_tests) %>% 
  select(city_state, estimate, conf.low, conf.high)
allcities_df

#Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

allcities_df %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

The original Washington Post dataset has records for 52,129 homicides (the rows) from 50 cities in the United States. The datset has 12 columns and includes variables such as victim_race, victim_last, victim_first, victim_age, victim_sex, reported_date and uid. 

Post data cleaning and creation of the city_state variable, there are only 2 variables and a total of 52,128 observations.The total number of homicides and the total number of uncilved homicides in each city_state is contained in the tital_hom data frame.

Baltimore has a sample estimate of 0.6455607. The 95% CI has an upper bound of 0.6276 and a lower bound of 0.6632.